Visualizing and Interpreting Multi-Group Confirmatory Factor Analysis

### 2023-03-31 manuscript.tex
update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

Enter author note here.

@@

Outline

\begin{itemize}
\item
  talk about LVM
\item
  talk about cfa
\item
  talk about mgcfa
\item
  how are measurement and replication/crisis related
\item
  why it is a bad idea to say that replication/crisis are \emph{because} bad measurement
\item
  how can we visualize and interpret MGCFA to help us understand the impact of measurement differences
\end{itemize}

By the end of this tutorial manuscript, readers will:

@@

The Brown (2015) steps of testing measurement invariance are demonstrated in this manuscript for illustration purposes, but in line with Stark, Chernyshenko, and Drasgow (2006) suggestions, the visualizations show the impact of loadings and intercepts together. The configural model was analyzed nesting both groups into the same CFA model requiring that both groups show the same model structure, but all other parameters are free to vary between groups. The metric model constrained the factor loadings of each group to be equal within the model. The scalar model then constrained the item intercepts (i.e., item mean) to be equal across groups. Finally, the strict model constrained the item variances (i.e., error variances) to be equal for each item across groups. These models are normally tested sequentially, and a convenience function \texttt{mgcfa} is provided in the supplemental documents for this manuscript.

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

Thank you to K.D. Valentine and Chelsea Parlett-Pelleriti for feedback on some ugly graphs.

@@

Psychological assessments play a critical role in our ability to measure and analyze constructs to support theories and experimental hypotheses. Defining and creating assessments to validly and reliability measure constructs is often difficult because phenomenon, such as anxiety, are often not directly observable. Instead, we use surveys and questionnaires to indirectly assess the underlying construct (DeVellis \& Thorpe, 2022). Latent variable modeling (i.e., structural equation modeling) is a popular tool for the validation of developed survey instruments to verify scale dimensionalality, structure, and model fit. A simple search for scale development reveals thousands of articles in psychology that examine new and previously published work, thus, illustrating the interest in both measurement and the use of validation techniques. Unfortunately, except in specialty journals, much of the validity evidence and/or development for measures used in empirical studies is not reported within the journal article Weidman, Steckler, \& Tracy (2017). Without this information, it is difficult to interpret individual study conclusions, as validity information allows for judgment of usefulness of the measured values (Flake \& Fried, 2020). Further, the current focus on replication Zwaan, Etz, Lucas, \& Donnellan (2018), reproducibility (Nelson et al., 2018), and the credibility of our results (Vazire, Schiavone, \& Bottesini, 2022) has demonstrated questionable measurement practices - decisions that researchers make like survey selection and scoring that impact the results of the study (Flake \& Fried, 2020). Transparent reporting of the use and creation of scales can improve both interpretation and reproducibility when using surveys developed to measure latent constructs (Shadish, Cook, \& Campbell, 2001).

A secondary concern for developed measures is the potential for differential responding and assessment within target populations. For example, Trent et al. (2013) examined for potential variability in the Revised Child Anxiety and Depression Scale in White and Black youths (Chorpita, Yim, Moffitt, Umemoto, \& Francis, 2000). They found that the scale mostly functioned the same for both White and Black individuals but differences in averages on individual items could potentially affect the scoring and interpretation of the scale results. This comparison of sub-populations is the test of measurement invariance (Meredith, 1993). Invariance or equivalence implies that the scale operates in the same fashion for each sub-group, and thus, differences in the final latent variable scores can interpretated as differences in populations. Non-invariance suggests that individuals respond or interpret items differently, and thus, differences in scores may represent different scores on the latent variable in the population or differences in measurement. Non-invariant measurement may lead to misleading results when making group comparisons, and assessing invariance has become a popular technique in scale development (Van De Schoot, Schmidt, De Beuckelaer, Lek, \& Zondervan-Zwijnenburg, 2015).

Measurement invariance is typically analyzed using confirmatory factor analysis, specifically, multi-group confirmatory factor analysis (MGCFA) or less often, with item response theory Tay, Meade, \& Cao (2015). First, the model is examined with the factor structure proposed for the latent and observed variables, and then often these models are assessed for each group separately. The two models are then combined together into one nested CFA in order to determine configural invariance Kline (2016). Configural invariance tests if the proposed factor structure is the same between groups. In this model, all other estimated parameters are allowed to vary between groups. The general approach is to use this model as a baseline for starting a sequential analysis of further restrictions between group parameters (i.e., more restrictive with each step). However, models without configural invariance can occur and often point to misspecification for the observed and latent variables within one group (i.e., cross loadings of items onto other latent variables or correlated error terms for one group only).

Next, the estimated parameter between each observed variable and its latent variable are constrained to be equal between groups for metric invariance. For example, item 1's factor loading must be equal to item 1's factor loading for each group. This test examines if the items represent the same relationship to the latent variable, or if specific items have weaker or stronger relationships in specific groups. Finding non-invariance at this stage generally points to items that have different functioning or interpretation for one group. At the third model, the item intercepts (i.e., item averages) are restricted across groups for scalar invariance. Scalar non-invariance would indicate that items have the same strength of relationship with their latent variable, just one group has a higher overall average on that item. Last (although sometimes not used), we may consider constraining error variances for each observed variable to be equal across groups for strict invariance. Strict non-invariance can occur when one group has a higher range of values on the observed variable, thus showing a larger variance. For example, if using a Likert scale, one group may use the full 1 to 7 range (creating a flatter distribution and larger variance), while the other group shows a ceiling effect of only using 5 to 7.

These concepts have been explored and implemented for the last fifty years Sörbom (1978) and implemented in the most popular structural equation modeling programs Boker et al. (2011). Byrne, Shavelson, and Muthén (1989) extended the ideas of multi-group testing by suggesting partial invariance (followed by Meredith, 1993). Partial invariance occurs when non-invariance is found but can be attributed to only a few parameter estimate differences between groups (i.e., items 1 and 2 have different factor loadings but all others are the same). This testing provided an advantage to understand where the potential non-invariance may occur for further study and interpretation guidelines. To determine when non-invariance and partial invariance occurred, each model is sequentially compared to the previous model using some form of a difference test. Traditionally, since models were nested, a chi-square difference test was used Cheung \& Rensvold (2002); however, given the known issues with chi-square (Thompson \& Daniel, 1996), people have favored empirical cutoffs for differences in fit indices. As the field pushes back against favoring cutoff criteria and rules of thumb Putnick \& Bornstein (2016), an effect size measure for translating ``how much'' non-invariance was developed \(d_{MACS}\) (Nye \& Drasgow, 2011). This effect size examines the differences in observed variables between the two groups for both the factor loading and the item intercept; thus, any differences in either or both will increase the effect size for non-invariance (Stark et al., 2006).

With \(d_{MACS}\) and measurement invariance testing, researchers can begin to quantify how and where their construct measurement may vary between groups. However, given the large number of studies that show non-invariance, it is clear that equivalence can be hard to meet. It is difficult to know if non-invariance occurs because of random sampling error, true population differences, or differences in replication and reproducibility of the construct in a new sample. Further, it is important to remember that the parameter estimates that we are testing are just that - estimates. All the parameter estimates have measures of standard error to indicate that they are more than likely variable with a new sample or population. Given that this information is generally ignored during the examination of measurement invariance, it may be that we are claiming that many scales are non-invariant, when in reality, the differences between loadings or item intercepts are small and unimportant. \(d_{MACS}\) provides the opportunity to begin to think about the smallest effect size of interest or the smallest meaningful effect size Anvari \& Lakens (2021). As mentioned, \(d_{MACS}\) has only really been explored for a combined intercept and loadings, and while useful, does not necessarily allow a researcher to pinpoint specific issues within an observed variable. The purpose of this manuscript is provide readers with a framework for visualization of differences in loadings, intercepts, and variances for each item, and the impact of those differences on the distribution of the latent mean. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researchers ability to judge the strength of the non-invariance differences and effect size for each item. Coupled with other indicators (i.e., fit indice differences, \(d_{MACS}\)), we can move toward a better understanding of how much measurement non-invariance is meaningful.

By the end of this manuscript, readers will:

@@

The Brown (2015) steps of testing measurement invariance are demonstrated in this manuscript for illustration purposes, but in line with Stark et al. (2006) suggestions, the visualizations show the impact of loadings and intercepts together. The configural model was analyzed nesting both groups into the same CFA model requiring that both groups show the same model structure, but all other parameters are free to vary between groups. The metric model constrained the factor loadings of each group to be equal within the model. The scalar model then constrained the item intercepts (i.e., item mean) to be equal across groups. Finally, the strict model constrained the item variances (i.e., error variances) to be equal for each item across groups. These models are normally tested sequentially, and a convenience function \texttt{mgcfa} is provided in the supplemental documents for this manuscript. Fit indices for the steps for multi-group models are presented in the appendix for comparison of cutoff rules of thumb (Cheung \& Rensvold, 2002) to effect sizes and visualizations presented in this manuscript. Fit indices include Akaike Information Criterion (AIC, Akaike, 1998), Bayesian Information Criterion (BIC, Schwarz, 1978), Comparative Fit Index (CFI, Bentler, 1990), Tucker Lewis Index (TLI, Tucker \& Lewis, 1973), root mean squared error of approximation RMSEA (Steiger, 1990), and standardized root mean squar residual (SRMR, Bentler, 1995).

@@

\hypertarget{code-examples}{%
\subsection{Code Examples}\label{code-examples}}

The complete code for this manuscript can be found at \url{https://osf.io/wev5f/}, and the function code for the convenience function for multi-group models and plots is found in the appendix. First, we would create our model code in \texttt{lavaan} syntax. The \texttt{lv} latent variable predicts the five measured variables, which are present as columns in our \texttt{df.invariant} dataset. You would include the dataframe in the \texttt{data} argument of our function, the name of the grouping variable in quotes for \texttt{group}, and the \texttt{lavaan} model syntax in the \texttt{model} argument. The \texttt{mgcfa} function code runs an overall model with all data, regardless of group, each group separately on the model, then the steps described above: configural, metric, scalar, and strict invariance.

\small

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# create lavaan model}
\NormalTok{model.overall }\OtherTok{\textless{}{-}} \StringTok{"lv =\textasciitilde{} q1 + q2 + q3 + q4 + q5}
\StringTok{q1 \textasciitilde{} 0*1}
\StringTok{lv \textasciitilde{} 1"}
\CommentTok{\# look at the data}
\FunctionTok{head}\NormalTok{(df.invariant)}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{verbatim}
##           q1          q2          q3         q4         q5   group
## 1 -0.8903542 -0.81707530  0.06137292 -1.3236407 -1.7916418 Group 1
## 2  1.1054521 -0.03540948 -0.81299606  1.0028340 -0.1909127 Group 1
## 3  1.4555852  1.54083484  1.59084213 -0.3345967 -0.6865496 Group 1
## 4 -1.8745187 -1.27880245 -2.53565792 -1.0024193 -1.6253249 Group 1
## 5 -0.4449517 -0.17782974  1.05507079 -1.2615705  1.7536428 Group 1
## 6  0.2278813  0.71348845  1.63251893  0.6449847 -1.0055700 Group 1
\end{verbatim}

\small

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# run our mgcfa function to run all models}
\NormalTok{results.invariant }\OtherTok{\textless{}{-}} \FunctionTok{mgcfa}\NormalTok{(}\AttributeTok{data =}\NormalTok{ df.invariant, }\CommentTok{\#dataframe}
                           \AttributeTok{group =} \StringTok{"group"}\NormalTok{, }
                           \AttributeTok{model =}\NormalTok{ model.overall)}
\CommentTok{\# what is saved for you}
\FunctionTok{names}\NormalTok{(results.invariant)}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{verbatim}
## [1] "model_coef"       "model_fit"        "model.overall"    "model.group1"    
## [5] "model.group2"     "model.configural" "model.metric"     "model.scalar"    
## [9] "model.strict"
\end{verbatim}

The results are saved as a list and include the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  \texttt{model\_coef}: a tidy dataframe with \emph{all} model's coefficients saved from the \texttt{lavaan} outputs.
\end{enumerate}

\small

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{head}\NormalTok{(results.invariant}\SpecialCharTok{$}\NormalTok{model\_coef)}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{verbatim}
## # A tibble: 6 x 13
##   term       op    estimate std.error statistic   p.value std.lv std.all std.nox
##   <chr>      <chr>    <dbl>     <dbl>     <dbl>     <dbl>  <dbl>   <dbl>   <dbl>
## 1 "lv =~ q1" =~       1        0          NA    NA         0.780   0.598   0.598
## 2 "lv =~ q2" =~       0.564    0.0864      6.52  6.99e-11  0.440   0.435   0.435
## 3 "lv =~ q3" =~       0.748    0.105       7.12  1.09e-12  0.583   0.505   0.505
## 4 "lv =~ q4" =~       0.338    0.0804      4.20  2.62e- 5  0.264   0.250   0.250
## 5 "lv =~ q5" =~       0.904    0.120       7.52  5.48e-14  0.705   0.613   0.613
## 6 "q1 ~1 "   ~1       0        0          NA    NA         0       0       0    
## # i 4 more variables: model <chr>, block <int>, group <int>, label <chr>
\end{verbatim}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  \texttt{model\_fit}: a tidy dataframe with \emph{all} model's fit indices saved from the \texttt{lavaan} outputs.
\end{enumerate}

\small

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{head}\NormalTok{(results.invariant}\SpecialCharTok{$}\NormalTok{model\_fit)}
\end{Highlighting}
\end{Shaded}

\normalsize

\begin{verbatim}
## # A tibble: 6 x 18
##    agfi   AIC   BIC   cfi chisq  npar  rmsea rmsea.conf.high   srmr   tli
##   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>           <dbl>  <dbl> <dbl>
## 1 0.979 7516. 7579. 0.994  6.37    15 0.0234          0.0697 0.0211 0.988
## 2 0.948 3766. 3819. 0.976  7.79    15 0.0473          0.108  0.0312 0.953
## 3 0.952 3762. 3815. 0.980  7.25    15 0.0424          0.104  0.0322 0.960
## 4 0.950 7528. 7654. 0.978 15.0     30 0.0449          0.0886 0.0317 0.956
## 5 0.942 7529. 7639. 0.954 24.7     26 0.0554          0.0905 0.0476 0.934
## 6 0.952 7523. 7616. 0.964 26.2     22 0.0428          0.0760 0.0488 0.960
## # i 8 more variables: converged <lgl>, estimator <chr>, ngroups <int>,
## #   missing_method <chr>, nobs <int>, norig <int>, nexcluded <int>, model <chr>
\end{verbatim}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  Saved \texttt{lavaan} fitted objects that you can use the \texttt{summary()}, \texttt{parameterEstimates()}, \texttt{fitIndices()}, etc. on. Overall model indicates the model without grouping variables testing all data on the proposed model structure. This model is then tested separately for each group (\texttt{model.group1}, \texttt{model.group2}). The final models follow the Brown (2015) naming convention for sequential steps for testing MGCFA for measurement invariance (\texttt{model.configural}, \texttt{model.metric}, \texttt{model.scalar}, \texttt{model.strict}).
\end{enumerate}

The results from the \texttt{model\_coef} table can then be used directly in the suggested plotting function. The plot outputs will be described below. First, here are the arguments for the function:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  \texttt{data\_coef}: A tidy dataframe of the parameter estimates from the models. This function assumes you have used \texttt{broom::tidy()} on the saved model from \texttt{lavaan} and added a column called ``model'' with the name of the model step. This function will only run for models that have used the grouping function (i.e., configural, metric, scalar, and strict or other combinations/steps you wish to examine).
\item
  \texttt{model\_step}: Which model do you want to plot? You should match this name to the one you want to extract from your model column in the \texttt{data\_coef}.
\item
  \texttt{item\_name}: Which observed variable from your model syntax do you want to plot? Please list this variable name exactly how it appears in the model.
\item
  \texttt{x\_limits}: What do you want the x-axits limits to be for your invariance plot? The default option is to assume the latent variable is standardized, and therefore, -1 to 1 is recommended. Use only two numbers, a lower and upper limit.
\item
  \texttt{y\_limits}: What do you want the y-axits limits to be for your invariance plot? Given that the latent variable is used to predict the observed values in the data, you could use the minimum and maximum values found in the data. If that range is large, consider reducing this value to be able to visualize the results (i.e., otherwise it may be too zoomed out to judge group differences). Use only two numbers, a lower and upper limit.
\item
  \texttt{ci\_level}: What confidence limit do you want to plot? Use 1 - \(\alpha\).
\item
  \texttt{model\_results}: In this argument, include the saved \texttt{lavaan} output for the model listed in the \texttt{model\_step} argument.
\item
  \texttt{lv\_name}: Include the name of the latent variable, exactly how it is listed in your \texttt{lavaan} syntax. You should plot the latent variable that the \texttt{item\_name} is linked to. If you have items that load onto multiple latent variables, you will need to make multiple plots.
\end{enumerate}

\small

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_mgcfa}\NormalTok{(}
  \AttributeTok{data\_coef =}\NormalTok{ results.invariant}\SpecialCharTok{$}\NormalTok{model\_coef, }\CommentTok{\# output from model\_coef}
  \AttributeTok{model\_step =} \StringTok{"configural"}\NormalTok{, }\CommentTok{\# which model do you want to plot}
  \AttributeTok{item\_name =} \StringTok{"q4"}\NormalTok{, }\CommentTok{\# name of observed item}
  \AttributeTok{x\_limits =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{), }\CommentTok{\# latent variable limits to graph}
  \AttributeTok{y\_limits =} \FunctionTok{c}\NormalTok{(}\FunctionTok{min}\NormalTok{(df.invariant}\SpecialCharTok{$}\NormalTok{q4), }\FunctionTok{max}\NormalTok{(df.invariant}\SpecialCharTok{$}\NormalTok{q4)), }\CommentTok{\# Y min and max in data }
  \AttributeTok{ci\_level =}\NormalTok{ .}\DecValTok{95}\NormalTok{, }\CommentTok{\# what ci do you want}
  \AttributeTok{model\_results =}\NormalTok{ results.invariant}\SpecialCharTok{$}\NormalTok{model.configural, }\CommentTok{\# what model results do you want }
  \AttributeTok{lv\_name =} \StringTok{"lv"} \CommentTok{\# which latent variable do you want }
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\normalsize

\hypertarget{visualization-of-invariance}{%
\subsection{Visualization of Invariance}\label{visualization-of-invariance}}

The output from this model can be found in Figure \ref{fig:invariant-pic}. On the left hand side, the item invariance is plotted, and on the right hand side, the latent mean distributions for the two groups are plotted. In the item invariance sub-plot, the visualization includes all three components traditionally seen in MGCFA testing steps: loadings, intercepts, and residuals. Each visualization element was designed to match the traditional visualization for that type of output. All parameter estimates are plotted on the unstandardized estimates and their confidence interval based on the standard error of the estimate.

\hypertarget{loadings}{%
\subsubsection{Loadings}\label{loadings}}

Factor loadings represent the slope of the regression equation for the latent variable predicting the scores on the observed variable (\(\hat{Y} \sim b_0 + b_1X + \epsilon\)). Therefore, the latent variable is shown on the x-axis using standardized values (i.e., \emph{z}-scores) where -1 indicates one standard deviation below the mean for the latent variable, 0 indicates the mean for the latent variable and so on. The y-axis indicates the observed variable scores, and here, the plot includes the entire range of the scale of the data for item four. The coefficient (\(b_1\)) for group 1 was 0.40, while the coefficient for group 2 was 0.34. The ribbon bands around the plotted slopes indicate the confidence interval for that estimate. In this plot, while the coefficients for each group are not literally equal, the overlapping and parallel slope bands indicate they are not different practically.

\hypertarget{intercepts}{%
\subsubsection{Intercepts}\label{intercepts}}

The item intercepts (\(b_0\)) are plotted on the middle line where they would cross the y-axis at a latent variable score of zero. These are represented by a dot with a set of confidence error bars around the point. The intercept for group 1 was 0.07, while the coefficient for group 2 was 0.03. In this invariant depiction, the overlap in the intercepts is clear, indicating they are not different. You can use \texttt{y\_limits} to zoom in on the graph if these are too small to be distinguishable.

\hypertarget{residuals}{%
\subsubsection{Residuals}\label{residuals}}

Residuals are trickier to plot, as they are the left over error when predicting the observed variables \(\epsilon\). It is tempting to plot this value as the confidence band around the slope, however, that defeats the purpose of understanding that the slopes are estimated separately from the residuals, and both have an associated variability around their parameter estimate. Therefore, residuals are represented in the inset picture at the bottom right of the item invariance plot. The black bars represent the estimated residual for each group (group 1: 0.91, group 2: 1.16). The distributions are plotted to represent the normal spread of values using the standard error of the residuals. The violin plot allows for direct comparison of those residuals and their potential distributions. Note that the placement has nothing to do with the x or y-axis and is designed to always show in the same location, regardless of size/value.

\hypertarget{latent-means}{%
\subsubsection{Latent Means}\label{latent-means}}

The overall impact of differences on the latent means can be found in the right hand visualization. The latent means are calculated by using the \texttt{predict} function and then plotted as overlapping histograms. The vertical colored lines represent the mean for each group, and the spread of the distribution can be examined using the density coloring. Finally, group labels are represented in the figure caption on the bottom. Group 1 is usually the group that is alphabetically first in the dataset or whichever group is the first that appears when using the \texttt{levels()} command.

@@

\hypertarget{graphing-effect-size}{%
\subsubsection{Graphing Effect Size}\label{graphing-effect-size}}
